|

### 2022-12-13 paper_comparison/paper_comparison/paper_comparison.Rmd
revised experiment section

The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference. Section \ref{experimental-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

  # Plot base data
  mod$plot(dat, theme = theme_light()) +
    ggtitle("A. Good residuals") + 
    xlab("fitted values") + 
    ylab("residuals"),

  # Include the z term
  # Reuse the x and e terms
  mod$plot(mod$
             set_prm("include_z", TRUE)$
             gen(300, computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("B. Non-linearity") +
    xlab("fitted values") + 
    ylab("residuals"),

  # Build a heter model
  # Reuse the x and e terms
  mod$plot(heter_model(b = 64)$
             gen(300, computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("C. Heteroskedasticity") +
    xlab("fitted values") + 
    ylab("residuals"),

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(He[2]:"U")) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[3]:"S")) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[6]:"M")) + 
  theme_light() -> p3

VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[18]:"Triple-U")) + 
  theme_light() -> p4

@@

Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [Hermite ref here] on random vector $\boldsymbol{x}$ formulated as

@@

```{r results='asis'}
data.frame(j = c("2", "3", "6", "18", ""),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", "NA"),
           sigma = c("0.25", "1.00", "2.00", "4.00", ""),
           a = c("-1", "0", "1", "", ""),
           b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
           n = c("50", "100", "300", "", "")) %>%
  knitr::kable(col.names = NULL,
               align = "rcrrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Description of all factors involved in the nonlinear and heteroskedasticity studies.', 
               label = "model-factor-table" ) %>%
  kableExtra::kable_styling(latex_options = "scale_down") %>%
  kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
                                 "Distribution of sym2\n", 
                                 "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
                                 "Poly Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
                                 "Heteroskedasticity (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
                                 "Size (sym6)")) %>% #"Sample size\n(sym6)")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym2", "$X_{raw}$", .) %>%
  gsub("sym3", "$\\\\sigma$", .) %>%
  gsub("sym4", "$a$", .) %>%
  gsub("sym5", "$b$", .) %>%
  gsub("sym6", "$n$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

@@

```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots. Four different distribution of $x_{raw}$ are used in the experiment to provide various visual patterns."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

VI_MODEL$plot(ori_dat) + 
  ggtitle("A. Uniform") + #quote(U(-1,1))) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle("B. Normal") + #quote(N(0,0.3^2))) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle("C. Skewed") + #quote(lognormal(0,0.6^2)/3)) + 
  theme_light() -> p3

VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle("D. Discrete") + #quote(U~"{1,5}")) + 
  theme_light() -> p4

@@

Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I. This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

@@

For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $x$, which will lead to the presence of heteroskedasticity. Visual patterns were simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I. A summary of the factors is given in Table \ref{tab:model-factor-table}.

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("A. a = -1: Left-triangle")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("B. a = 0: Butterfly")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("C. a = 1: Right-triangle")) +
  theme_light() -> p3

@@

```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["heter_169"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$. The actual data plot location was 15. 8 out of 11 subjects correctly identified the actual data plot for this lineup.

## Experimental setup

@@

As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters. A large value of $\sigma$ will increase the variation of the error of the non-linear model and decrease the visibility of the visual pattern. The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor. Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(n~"="~300)) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
  ggtitle(quote(n~"="~50)) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
  ggtitle(quote(n~"="~100)) + 
  theme_light() -> p3

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(sigma~"="~0.5)) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~1)) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~2)) + 
  theme_light() -> p3

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~4)) + 
  theme_light() -> p4

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("b = 0.25")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 1")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 4")) +
  theme_light() -> p3

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 16")) +
  theme_light() -> p4

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 64")) +
  theme_light() -> p5

@@

Three replications are made for each of the parameter values shown in Table \ref{tab:parameter-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups. In addition, each lineup is designed to be evaluated by five different subjects. After attempting some pilot studies internally in our research group, we decide to present a block of 20 lineups to every subject. And to ensure the quality of the survey data, two lineups with obvious visual patterns are included as attention checks. Thus, $576 \times 5/(20-2) = 160$ and $540 \times 5 /(20-2) = 150$ subjects were recruited to satisfy the design of the experiment I and experiment II respectively.

As mentioned in Section [the alpha section], $\alpha$ used in Equation \ref{eq:eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups. Hence, 36 Rorschach lineups with all combinations of $n$ and $X_{raw}$ and three replications are designed to be included in experiment III. All Rorschach lineups are planned to be evaluated by 20 subjects. However, presenting only Rorschach lineups to subjects are considered to be bad practices as subjects will lose interest quickly. We planned to also collect 6 more evaluations on the 279 lineups with uniform distribution, resulted in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3)/(20-2) = 133$ subjects recruited for experiment III.

@@

Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [ref here]. Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions. During the experiment, every subject was presented with a block of 20 lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. An additional horizontal red line was added at $y = 0$ as a helping line. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others. If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason. No subject was shown the same lineup twice. Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected. A subject's submission was only accepted if the actual data plot was identified for at least one attention check. Data of rejected submissions were discarded automatically to maintain the overall data quality.

>>>>

The persistence of this advice to check the plots is curious, and investigating why this might be common advice is the subject of this paper. The paper is structured as follows. The next background section describes the types of departures that one expects to detect, and describes a formal statistical process for reading residual plots, called visual inference. Section \ref{experiment-design} describes the experiment design to compare the decision made by formal hypothesis testing, and how humans would read diagnostic plots. The results are reported in Section \ref{results}. We finish with a discussion on future work, in particular how the responsibility for residual plot reading might be passed on to computer vision.

@@

  # Plot base data
  mod$plot(dat, theme = theme_light()) +
    ggtitle("A. Good residuals") + 
    xlab("Fitted values") + 
    ylab("Residuals"),

  # Include the z term
  # Reuse the x and e terms
  mod$plot(mod$
             set_prm("include_z", TRUE)$
             gen(300, computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("B. Non-linearity") +
    xlab("Fitted values") + 
    ylab("Residuals"),

  # Build a heter model
  # Reuse the x and e terms
  mod$plot(heter_model(b = 64)$
             gen(300, computed = select(dat, x, e)), 
    theme = theme_light()) + 
    ggtitle("C. Heteroskedasticity") +
    xlab("Fitted values") + 
    ylab("Residuals"),

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(He[2]:"U")) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[3]:"S")) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p2

VI_MODEL$plot(poly_model(shape = 3, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[6]:"M")) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p3

VI_MODEL$plot(poly_model(shape = 4, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.05)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote(He[18]:"Triple-U")) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p4

@@

Experiment I is designed to study the ability of human subjects to detect the effect of a non-linear term $\boldsymbol{z}$ constructed using Hermite polynomials [unsure how to cite 1810 work] on random vector $\boldsymbol{x}$ formulated as

@@

```{r results='asis'}
data.frame(j = c("2", "3", "6", "18", ""),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$", ""),
           sigma = c("0.25", "1.00", "2.00", "4.00", ""),
           a = c("-1", "0", "1", "", ""),
           b = c("0.25", "1.00", "4.00", "16.00", "64.00"),
           n = c("50", "100", "300", "", "")) %>%
  knitr::kable(col.names = NULL,
               align = "rcrrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Description of all factors involved in the non-linear and heteroskedasticity studies.', 
               label = "model-factor-table" ) %>%
  kableExtra::kable_styling(latex_options = "scale_down") %>%
  kableExtra::add_header_above(c("Poly Order (sym1)", #"Order of Hermite polynomial\n(sym1)", 
                                 "Distribution of sym2\n", 
                                 "SD (sym3)", #"Standard deviation of polynomial model\n(sym3)",
                                 "Heteroskedasticity Shape (sym4)", # "Shape of heteroskedasticity\n(sym4)", 
                                 "Heteroskedasticity (sym5)", # Variance factor of heteroskedasticity model\n(sym5)",
                                 "Size (sym6)")) %>% #"Sample size\n(sym6)")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym2", "$X_{raw}$", .) %>%
  gsub("sym3", "$\\\\sigma$", .) %>%
  gsub("sym4", "$a$", .) %>%
  gsub("sym5", "$b$", .) %>%
  gsub("sym6", "$n$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

@@

```{r different-dist, fig.height = 2, fig.cap="Variations in fitted values, that might affect perception of residual plots. Four different distribution of $X_{raw}$ are used in the experiment to provide various visual patterns."}
ori_dat <- poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300)

VI_MODEL$plot(ori_dat) + 
  ggtitle("A. Uniform") + #quote(U(-1,1))) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p1

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle("B. Normal") + #quote(N(0,0.3^2))) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p2

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_lognormal(sigma = 0.6); closed_form(~stand_dist(raw_x/3 - 1))}, sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle("C. Skewed") + #quote(lognormal(0,0.6^2)/3)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p3

VI_MODEL$plot(poly_model(shape = 1, x = rand_uniform_d(k = 5, even = TRUE), sigma = 0.5)$gen(300, computed = select(ori_dat, e))) +
  ggtitle("D. Discrete") + #quote(U~"{1,5}")) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p4

@@

Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I. This lineup is produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location is $2^3 - 4$. All five subjects correctly identified the actual data plot for this lineup.

@@

For $b \neq 0$, the variance-covariance matrix of the error term $\boldsymbol{\varepsilon}$ is correlated with the regressor $\boldsymbol{x}$, which will lead to the presence of heteroskedasticity. Visual patterns were simulated using three different shapes ($a$ = -1, 0, 1) and the same four different distribution of $X_{raw}$ used in experiment I. A summary of the factors is given in Table \ref{tab:model-factor-table}.

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("A. a = -1: Left-triangle")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("B. a = 0: Butterfly")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p2

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("C. a = 1: Right-triangle")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p3

@@

```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}

# vi_survey %>%
#   filter(!attention_check) %>%
#   filter(!null_lineup) %>%
#   filter(type == "heteroskedasticity") %>%
#   filter(x_dist == "uniform") %>%
#   filter(a == -1) %>%
#   filter(n == 300) %>%
#   group_by(unique_lineup_id) %>%
#   summarise(effect_size = first(effect_size)) %>%
#   arrange(effect_size)

# vi_survey %>%
#   filter(unique_lineup_id == "heter_471") %>%
#   View()

VI_MODEL$plot_lineup(vi_lineup$heter_471$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$. The actual data plot location is $2^4 + 2$. Nine out of 11 subjects correctly identified the actual data plot for this lineup.

## Experiment setup

@@

As summarised in Table \ref{tab:model-factor-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. Parameter $\sigma \in \{0.5, 1, 2, 4\}$ and $b \in \{0.25, 1, 4, 16, 64\}$ were used in experiment I and II respectively. Figure \ref{fig:different-sigma} and \ref{fig:different-b} demonstrate the impact of these two parameters. A large value of $\sigma$ will increase the variation of the error of the non-linearity model and decrease the visibility of the visual pattern. The parameter $b$ controls the ratio of the standard deviation of the heteroskedasticity across the domain of the regressor. Given $x \neq a$, a larger value of $b$ will lead to a larger ratio of the variance at $x$ to the variance at $x - a = 0$, making the visual pattern more obvious.

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(n~"="~300)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
  ggtitle(quote(n~"="~50)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p2

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
  ggtitle(quote(n~"="~100)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p3

@@

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(sigma~"="~0.5)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p1

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~1)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p2

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 2)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~2)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p3

VI_MODEL$plot(poly_model(shape = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 4)$gen(300, computed = select(ori_dat, x))) +
  ggtitle(quote(sigma~"="~4)) + 
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p4

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("b = 0.25")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 1)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 1")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p2

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 4)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 4")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p3

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 16)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 16")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p4

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 64)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("b = 64")) +
  theme_light() +
  xlab("Fitted values") +
  ylab("Residuals") -> p5

@@

Three replications were made for each of the parameter values shown in Table \ref{tab:model-factor-table} resulting in $(4 \times 4 \times 4 \times 3 + 4 \times 3 \times 5 \times 3) \times 3 = 1116$ different lineups. In addition, each lineup was designed to be evaluated by five different subjects. After attempting some pilot studies internally in our research group, we decided to present a block of 20 lineups to every subject. And to ensure the quality of the survey data, two lineups with obvious visual patterns were included as attention checks. Thus, $576 \times 5 \div (20-2) = 160$ and $540 \times 5 \div (20-2) = 150$ subjects were recruited to satisfy the design of the experiment I and experiment II respectively.

As mentioned in Section \ref{calculating-p-values}, $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} needs to be estimated using Rorschach lineups. Hence, 36 lineups consist of only null plots with all combinations of $n$ and $X_{raw}$ and three replications are designed to be included in experiment III. Note that we generated the lineups using the same simulation mechanism discussed in Section \ref{sampling-from-the-null-distribution} rather than from a null model. We overlooked this detail until the data had been collected. Thus, we used another valid method suggested in vanderplas2021statistical which was designed for typical lineups to estimate $\alpha$ correctly. However, by not using Rorschach lineups, we did lose some accuracy in the estimation. We recommend future studies to generate Rorschach lineups strictly following the definition given by buja_statistical_2009.

All lineups consist of only null plots are planned to be evaluated by 20 subjects. However, presenting only these lineups to subjects are considered to be bad practices as subjects will lose interest quickly. We planned to also collect 6 more evaluations on the 279 lineups with $X_{raw} \sim U(-1,1)$, resulted in $(36 \times 20 + (4 \times 4 \times 3 + 3 \times 5 \times 3) \times 3) \div (20-2) = 133$ subjects recruited for experiment III.

@@

Subjects for all three experiments were recruited from an crowdsourcing platform called Prolific [palan2018prolific]. Prescreening procedure was applied during the recruitment, subjects were required to be fluent in English, with $98\%$ minimum approval rate in other studies and 10 minimum submissions. During the experiment, every subject was presented with a block of 20 lineups. For each lineup, the actual data plot was drawn as a standard residual plot of the null model with raw residuals on the y-axis and fitted values on the x-axis. An additional horizontal red line was added at $y = 0$ as a helping line. The 19 null datasets were generated by the residual rotation technique, and plotted in the same way. The lineup consisted of 20 residual plots with one randomly placed actual data plot. And for every lineup, the subject was asked to select one or more plots that are most different from others, provide a reason for their selections, and evaluate how different they think the selected plots were from others. If there was no noticeable difference between plots in a lineup, subjects were permitted to select zero plots without providing the reason. No subject was shown the same lineup twice. Information about preferred pronoun, age group, education, and previous experience in visual experiment were also collected. A subject's submission was only accepted if the actual data plot was identified for at least one attention check. Data of rejected submissions were discarded automatically to maintain the overall data quality.
